Introducing STRAIN: A simple and interpretable pass rush metric inspired by materials science

### 2023-03-10 paper/paper.Rmd
built figures + pipeline

@@

@@

@@

@@

>>>>

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
library(here)
```

@@

```{r}
#| label: fig_field
#| fig.cap: "snapshots"
#| fig.width: 7
#| fig.height: 9
source(here("scripts", "fig_field.R"))
fig_field
```

Figure \ref{fig:fig_field} shows snapshots

```{r}
#| label: fig_crosby_curves
#| fig.cap: "crosby"
#| fig.width: 5.1
#| fig.height: 3
source(here("scripts", "fig_crosby_curves.R"))
fig_crosby_curves
```

Figure \ref{fig:fig_crosby_curves} shows Crosby throughout the play

@@

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions"
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "fig_pos_curves.R"))
fig_pos_curves
```

Figure \ref{fig:fig_pos_curves} shows pos curves

@@

```{r}
#| label: fig_cor_pressure
#| fig.cap: "Cor with"
#| fig.width: 4.2
#| fig.height: 3
source(here("scripts", "fig_cor_pressure.R"))
fig_cor_pressure
```

Figure \ref{fig:fig_cor_pressure} shows cor with pressure

Fairly strong correlation (r = 0.6254943) between average STRAIN & hurries + sacks + hits per snaps

@@

```{r}
#| label: fig_stability
#| fig.cap: "Metric stability"
#| fig.width: 4.2
#| fig.height: 3
source(here("scripts", "fig_stability.R"))
fig_stability
```

0.8544865

Figure \ref{fig:fig_stability} shows

Average STRAIN across all frames played in first and last 4 weeks

Strong correlation shows great metric stability

### 2023-03-10 paper/paper.tex
built figures + pipeline

---
title: |
  Introducing STRAIN: A simple and interpretable pass rush metric inspired by materials science
short_title: STRAIN
authors: 
- name: "Author 1"
  affiliation: "Address"
  city: "City"
  country: "Country"
  email: "email1\\@example.com"
- name: "Author 2"
  affiliation: "Address2"
  city: "City2"
  country: "Country2"
  email: "email2\\@example.com"
abstract: |
  Abstract goes here
keywords: tracking data, Bornn legacy, too much Canadian
bibliography: "bibliography"
output: rticles::rss_article
header-includes:
  \usepackage{setspace}
  \usepackage{booktabs}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
library(here)
```

@@

# Introduction

<!--
TODO: update citations after they're published
-->

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state-of-the-art tool in sports analytics.
Numerous sports have been collecting and releasing data on player and ball locations on the playing surface over the course of a game.
This spatio-temporal multi-resolution source of data have provided exceptional opportunities for researchers to perform advanced studies at a more complex level to deepen the understanding of different sports.
For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

The National Football League (NFL) introduced their player tracking system known as Next Gen Stats (NGS) in 2016
BDB
https://operations.nfl.com/gameday/technology/nfl-next-gen-stats/

@@

## Previous Pass Rush Metrics

@@

https://www.pro-football-reference.com/about/advanced_stats.htm

Pressures: QB pressures (hurries + knockdowns + sack plays). Note that "sack plays" in this instance include both sacks and assists.

@@

https://www.pff.com/news/pro-signature-stats-spotlight-defensive-line

@@

@Burke2018Created

https://www.espn.com/nfl/story/_/id/24892208

@@

## Contributions of STRAIN

@@

# Methods

## Data

2023 bdb [@Howard2023NFL]

@@

## Strain rate in materials science

## Application to pass-rushing in football

# Results

## Example play

```{r}
#| label: fig_field
#| fig.cap: "snapshots"
#| fig.width: 7
#| fig.height: 9
source(here("scripts", "fig_field.R"))
fig_field
```

@@

```{r}
#| label: fig_crosby_curves
#| fig.cap: "crosby"
#| fig.width: 5.1
#| fig.height: 3
source(here("scripts", "fig_crosby_curves.R"))
fig_crosby_curves
```

@@

## Positional STRAIN curves

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions"
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "fig_pos_curves.R"))
fig_pos_curves
```

@@

## Ranking the best pass-rushers

## Corr with pressure rate = (hits + sacks + hurries) per snap

```{r}
#| label: fig_cor_pressure
#| fig.cap: "Cor with"
#| fig.width: 4.2
#| fig.height: 3
source(here("scripts", "fig_cor_pressure.R"))
fig_cor_pressure
```

@@

Fairly strong correlation (r = 0.6254943) between average STRAIN & hurries + sacks + hits per snaps

## Statistical Properties of STRAIN

From @Franks2016Meta (github https://github.com/afranks86/meta-analytics)

@@

```{r}
#| label: fig_stability
#| fig.cap: "Metric stability"
#| fig.width: 4.2
#| fig.height: 3
source(here("scripts", "fig_stability.R"))
fig_stability
```

@@

# Discussion

# References

>>>>

\documentclass{statsoc}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{amssymb, amsmath, geometry}
\usepackage{natbib}
\usepackage{hyperref}

\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\title[STRAIN]{Introducing STRAIN: A simple and interpretable pass rush
metric inspired by materials science}

\author[]{Author 1}
\address{Address,
City,
Country}
\email{email1@example.com}

\author[]{Author 2}
\address{Address2,
City2,
Country2}
\email{email2@example.com}

% BIBLIOGRAPHY
\usepackage[authoryear]{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{setspace} \usepackage{booktabs}

\begin{document}

\begin{abstract}
Abstract goes here
\end{abstract}
\keywords{tracking data, Bornn legacy, too much Canadian}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In recent years, tracking data have replaced traditional box-score
statistics and play-by-play data as the state-of-the-art tool in sports
analytics. Numerous sports have been collecting and releasing data on
player and ball locations on the playing surface over the course of a
game. This spatio-temporal multi-resolution source of data have provided
exceptional opportunities for researchers to perform advanced studies at
a more complex level to deepen the understanding of different sports.
For complete surveys on how tracking data have transformed sports
analytics, see \citet{Macdonald2020Recreating}, \citet{Baumer2023Big},
and \citet{Kovalchik2023Player}.

The National Football League (NFL) introduced their player tracking
system known as Next Gen Stats (NGS) in 2016 BDB
\url{https://operations.nfl.com/gameday/technology/nfl-next-gen-stats/}

@@

\hypertarget{previous-pass-rush-metrics}{%
\subsection{Previous Pass Rush
Metrics}\label{previous-pass-rush-metrics}}

@@

\url{https://www.pro-football-reference.com/about/advanced_stats.htm}

Pressures: QB pressures (hurries + knockdowns + sack plays). Note that
``sack plays'' in this instance include both sacks and assists.

@@

\url{https://www.pff.com/news/pro-signature-stats-spotlight-defensive-line}

@@

\citet{Burke2018Created}

\url{https://www.espn.com/nfl/story/_/id/24892208}

@@

\hypertarget{contributions-of-strain}{%
\subsection{Contributions of STRAIN}\label{contributions-of-strain}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

2023 bdb \citep{Howard2023NFL}

@@

\hypertarget{strain-rate-in-materials-science}{%
\subsection{Strain rate in materials
science}\label{strain-rate-in-materials-science}}

\hypertarget{application-to-pass-rushing-in-football}{%
\subsection{Application to pass-rushing in
football}\label{application-to-pass-rushing-in-football}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{example-play}{%
\subsection{Example play}\label{example-play}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig_field-1}

}

\caption{snapshots}\label{fig:fig_field}
\end{figure}

@@

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig_crosby_curves-1}

}

\caption{crosby}\label{fig:fig_crosby_curves}
\end{figure}

@@

\hypertarget{positional-strain-curves}{%
\subsection{Positional STRAIN curves}\label{positional-strain-curves}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig_pos_curves-1}

}

\caption{STRAIN curves for different positions}\label{fig:fig_pos_curves}
\end{figure}

@@

\hypertarget{ranking-the-best-pass-rushers}{%
\subsection{Ranking the best
pass-rushers}\label{ranking-the-best-pass-rushers}}

\hypertarget{corr-with-pressure-rate-hits-sacks-hurries-per-snap}{%
\subsection{Corr with pressure rate = (hits + sacks + hurries) per
snap}\label{corr-with-pressure-rate-hits-sacks-hurries-per-snap}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig_cor_pressure-1}

}

\caption{Cor with}\label{fig:fig_cor_pressure}
\end{figure}

@@

Fairly strong correlation (r = 0.6254943) between average STRAIN \&
hurries + sacks + hits per snaps

\hypertarget{statistical-properties-of-strain}{%
\subsection{Statistical Properties of
STRAIN}\label{statistical-properties-of-strain}}

From \citet{Franks2016Meta} (github
\url{https://github.com/afranks86/meta-analytics})

@@

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig_stability-1}

}

\caption{Metric stability}\label{fig:fig_stability}
\end{figure}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\bibliographystyle{rss}
\bibliography{bibliography}

\end{document}
